Contextuality and the Born rule in fuzzy quantum theories

### 2017-04-24 prop_iop.tex
prop_iop.tex \section{Expectation Values for IVPMs} Keep proving the details...

\begin{thm}\label{thm:sub-event-space}Given a Hilbert space~$\Hilb$
and a mutually commuting sub-event space~$\events'\subseteq\events$,
there is set~$\Omega$ and a function~$\tau:\events'\rightarrow2^{\Omega}$
such that:
\begin{itemize}
\item If $\mu:\events'\rightarrow\left[0,1\right]$ is a quantum real-valued
probability measure, then there is a classical real-valued probability
measure~$\mu':\Omega\rightarrow\left[0,1\right]$ such that $\mu\left(P\right)=\mu'\left(\tau\left(P\right)\right)$
for all $P\in\events'$.
\item If $\bar{\mu}:\events'\rightarrow\mathscr{I}$ is a quantum interval-valued
probability measure, then there is a classical interval-valued probability
measure~$\bar{\mu}':\Omega\rightarrow\mathscr{I}$ such that $\bar{\mu}\left(P\right)=\bar{\mu}'\left(\tau\left(P\right)\right)$
for all $P\in\events'$
\end{itemize}
\end{thm}

@@

\begin{lemma}Follow the same notation in the definition, we also
define $X^{+}=\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}$
and $X^{-}=\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}$. Then,
\begin{equation}
\int X\rmd\bar{\mu}=\int X^{+}\rmd\bar{\mu}-\int\left(-X^{-}\right)\rmd\bar{\mu}\textrm{ .}
\end{equation}

\end{lemma}

@@

\begin{thm}\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}For
every classical IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ and
any random variable~$X:\Omega\rightarrow\mathbb{R}$, $\bar{\mu}$
is convex if and only if 
\begin{equation}
\int X\rmd\bar{\mu}=\left[\min_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu,\max_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu\right]\textrm{ .}
\end{equation}
\end{thm}

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\mul\left(E\right),\mur\left(E\right)\right]$.
This time, when we express the observable~$\mathbf{O}$ as its spectrum
decomposition~$\sum_{i=1}^{N^{-}}\lambda_{i}^{-}P_{i}^{-}+\sum_{i=1}^{N^{+}}\lambda_{i}^{+}P_{i}^{+}$,
we separate the positive and negative eigenvalues and order them from
the smallest to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}\le\ldots\le\lambda_{1}^{-}\le0\le\lambda_{1}^{+}\le\ldots\le\lambda_{N^{+}}^{+}\textrm{ .}
\end{equation}
Let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$ and $\Delta\mu^{*}\left(P_{i}^{s\prime}\right)=\mu^{*}\left(P_{i}^{s\prime}\right)-\mu^{*}\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
We can combine the classical Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with Eq.~(\ref{eq:quantum-expectation}) to define the expectation
value of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta\mul\left(P_{i}^{+\prime}\right),\Delta\mur\left(P_{i}^{+\prime}\right)\right]+\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta\mul\left(P_{i}^{-\prime}\right),\Delta\mur\left(P_{i}^{-\prime}\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{equation}
Let $\events\left(\mathbf{O}\right)$ denote the minimal sub-event
space containing $\left\{ P_{i}\right\} _{i=1}^{N}$. Then, we have
the corresponding theorem.

\begin{thm}Given a Hilbert space of $d\ge3$, for every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
and any observable~$\mathbf{O}$, if $\bar{\mu}$ is convex, then
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events\left(\mathbf{O}\right)\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events\left(\mathbf{O}\right)\right)}\expval{\mathbf{O}}_{\rho}\right]\textrm{ .}\label{eq:quantum-interval-expectation-core}
\end{equation}
Conversely, if equation~(\ref{eq:quantum-interval-expectation-core})
holds for every observable, then $\bar{\mu}$ is convex.\end{thm}

>>>>

\begin{thm}\label{thm:sub-event-space}Given a Hilbert space~$\Hilb$
and a mutually commuting sub-event space~$\events'\subseteq\events$,
there is set~$\Omega$ and a function~$\tau:\events'\rightarrow2^{\Omega}$
such that:
\begin{itemize}
\item If $\mu:\events'\rightarrow\left[0,1\right]$ is a quantum real-valued
probability measure, then there is a classical real-valued probability
measure~$\mu':2^{\Omega}\rightarrow\left[0,1\right]$ such that $\mu\left(P\right)=\mu'\left(\tau\left(P\right)\right)$
for all $P\in\events'$.
\item If $\bar{\mu}:\events'\rightarrow\mathscr{I}$ is a quantum interval-valued
probability measure, then there is a classical interval-valued probability
measure~$\bar{\mu}':2^{\Omega}\rightarrow\mathscr{I}$ such that
$\bar{\mu}\left(P\right)=\bar{\mu}'\left(\tau\left(P\right)\right)$
for all $P\in\events'$
\end{itemize}
\end{thm}

@@

\begin{lemma}Follow the same notation in the definition, we also
define $X^{+}=\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}$
and $X^{-}=\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}$. Then,
$\int X\rmd\bar{\mu}=\int X^{+}\rmd\bar{\mu}-\int\left(-X^{-}\right)\rmd\bar{\mu}$.\end{lemma}

@@

\begin{thm}\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}For
every classical IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ and
any random variable~$X:\Omega\rightarrow\mathbb{R}$, $\bar{\mu}$
is convex if and only if $\int X\rmd\bar{\mu}=\left[\min_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu,\max_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu\right]$.\end{thm}

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(P\right)=\left[\mul\left(P\right),\mur\left(P\right)\right]$
for all $P\in\events$. This time, when we express the observable
as its spectrum decomposition
\begin{equation}
\mathbf{O}=\sum_{i=1}^{N^{-}}\lambda_{i}^{-}P_{i}^{-}+\sum_{i=1}^{N^{+}}\lambda_{i}^{+}P_{i}^{+}\textrm{ ,}\label{eq:spectrum-decomposition}
\end{equation}
we separate the positive and negative eigenvalues and order its eigenvalues
from the smallest to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}\le\ldots\le\lambda_{1}^{-}\le0\le\lambda_{1}^{+}\le\ldots\le\lambda_{N^{+}}^{+}\textrm{ .}
\end{equation}
Let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$ and $\Delta\mu^{*}\left(P_{i}^{s\prime}\right)=\mu^{*}\left(P_{i}^{s\prime}\right)-\mu^{*}\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
We can combine the classical Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with Eq.~(\ref{eq:quantum-expectation}) and define the expectation
value of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta\mul\left(P_{i}^{+\prime}\right),\Delta\mur\left(P_{i}^{+\prime}\right)\right]+\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta\mul\left(P_{i}^{-\prime}\right),\Delta\mur\left(P_{i}^{-\prime}\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{equation}

Intuitively, the expectation value of an observable with respect to
an IVPM should be the minimum and maximum expectation value of the
same observable with respect to its consistent states. We found this
is true when we require our IVPM to be convex, and on a commuting
sub-event space, whose definition is given as follow.

\begin{definition}A state~$\rho$ is consistent with $\bar{\mu}$
on a projector~$P$ if $\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)$,
and denoted by $\rho\in\coreBorn\left(\bar{\mu},\left\{ P\right\} \right)$.\end{definition}

\begin{definition}Given an observable~$\mathbf{O}$ with its spectrum
decomposition~(\ref{eq:spectrum-decomposition}). A sub-event space~$\events'\subseteq\events$
is called a sub-event space induced by $\mathbf{O}$ if $\left\{ P_{i}^{-}\right\} _{i=1}^{N^{-}}\cup\left\{ P_{i}^{+}\right\} _{i=1}^{N^{+}}\subseteq\events'$.

\end{definition}

\noindent Having these two definition, we can write down our idea
formally as follow.

\begin{thm}Given a Hilbert space of $d\ge3$, for every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
and any observable~$\mathbf{O}$ inducing a sub-event space~$\events'$,
if $\bar{\mu}$ is convex, then
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\textrm{ .}\label{eq:quantum-interval-expectation-core}
\end{equation}
Conversely, if Eq.~(\ref{eq:quantum-interval-expectation-core})
holds for every observable, then $\bar{\mu}$ is convex.\end{thm}

\begin{proof}Since $\events'$ is induced by $\mathbf{O}$, it is
commuting. By theorem~\ref{thm:sub-event-space}, there is set~$\Omega$,
a function~$\tau:\events'\rightarrow2^{\Omega}$, and a convex classical
interval-valued probability measure~$\bar{\mu}':2^{\Omega}\rightarrow\mathscr{I}$
such that $\bar{\mu}=\bar{\mu}'\circ\tau$. Let $\mul[\prime]:\events'\rightarrow\left[0,1\right]$
and $\mur[\prime]:\events'\rightarrow\left[0,1\right]$ are the left-end
and the right-end of $\bar{\mu}'$, respectively, i.e., $\bar{\mu}'\left(E\right)=\left[\mul[\prime]\left(E\right),\mur[\prime]\left(E\right)\right]$
for all $E\in2^{\Omega}$. Consider the random variable~$X:\Omega\rightarrow\mathscr{I}$
defined by $X=\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\mathbf{1}_{\tau\left(P_{i}^{-}\right)}+\sum_{i=1}^{N^{+}}\lambda_{i}^{+}\mathbf{1}_{\tau\left(P_{i}^{+}\right)}$.
Then, 
\begin{eqnarray*}
\expval{\mathbf{O}}_{\bar{\mu}} & = & \sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta\mul\left(P_{i}^{+\prime}\right),\Delta\mur\left(P_{i}^{+\prime}\right)\right]+\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta\mul\left(P_{i}^{-\prime}\right),\Delta\mur\left(P_{i}^{-\prime}\right)\right]\\
 & = & \sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta\mul[\prime]\left(\tau\left(P_{i}^{+\prime}\right)\right),\Delta\mur[\prime]\left(\tau\left(P_{i}^{+\prime}\right)\right)\right]+\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta\mul[\prime]\left(\tau\left(P_{i}^{-\prime}\right)\right),\Delta\mur[\prime]\left(\tau\left(P_{i}^{-\prime}\right)\right)\right]\\
 & = & \int X\rmd\bar{\mu}'\\
 & = & \left[\min_{\mu\in\mathrm{core}\left(\bar{\mu}'\right)}\int X\rmd\mu,\max_{\mu\in\mathrm{core}\left(\bar{\mu}'\right)}\int X\rmd\mu\right]
\end{eqnarray*}
\end{proof}

### 2017-04-24 prop_letter.tex
prop_iop.tex \section{Expectation Values for IVPMs} Keep proving the details...

Formally, we can ask: what can we deduce about the state of a quantum
system given a quantum interval-valued probability measure, i.e.,
given observations done with finite resources. Since the probabilities
are now intervals, it is intuitive to say a state~$\rho$ is consistent
with $\bar{\mu}$ on a projector~$P$ if 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)\textrm{ ,}
\end{equation}
and denoted by $\rho\in\coreBorn\left(\bar{\mu},\left\{ P\right\} \right)$.

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\mul\left(E\right),\mur\left(E\right)\right]$
for all $E\in\events$. This time, when we express the observable
as its spectrum decomposition
\begin{equation}
\mathbf{O}=\sum_{i=1}^{N^{-}}\lambda_{i}^{-}P_{i}^{-}+\sum_{i=1}^{N^{+}}\lambda_{i}^{+}P_{i}^{+}\textrm{ ,}\label{eq:spectrum-decomposition}
\end{equation}
we separate the positive and negative eigenvalues and order its eigenvalues
from the smallest to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}\le\ldots\le\lambda_{1}^{-}\le0\le\lambda_{1}^{+}\le\ldots\le\lambda_{N^{+}}^{+}\textrm{ .}
\end{equation}
Let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$ and $\Delta\mu^{*}\left(P_{i}^{s\prime}\right)=\mu^{*}\left(P_{i}^{s\prime}\right)-\mu^{*}\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
We can combine the classical Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with Eq.~(\ref{eq:quantum-expectation}) and define the expectation
value of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{eqnarray}
\expval{\mathbf{O}}_{\bar{\mu}} & = & \sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta\mul\left(P_{i}^{+\prime}\right),\Delta\mur\left(P_{i}^{+\prime}\right)\right]\nonumber \\
 &  & +\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta\mul\left(P_{i}^{-\prime}\right),\Delta\mur\left(P_{i}^{-\prime}\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{eqnarray}

@@

\noindent Having the above two definitions, we can write down our
idea formally as follow.

>>>>

Formally, we can ask: what can we deduce about the state of a quantum
system given a quantum interval-valued probability measure, i.e.,
given observations done with finite resources. Since the probabilities
are now intervals, it is intuitive to say a state~$\rho$ is consistent
with $\bar{\mu}$ on a projector~$P$ if $\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)$,
and denoted by $\rho\in\coreBorn\left(\bar{\mu},\left\{ P\right\} \right)$.

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(P\right)=\left[\mul\left(P\right),\mur\left(P\right)\right]$
for all $P\in\events$. This time, when we express the observable
as its spectrum decomposition
\begin{equation}
\mathbf{O}=\sum_{i=1}^{N^{-}}\lambda_{i}^{-}P_{i}^{-}+\sum_{i=1}^{N^{+}}\lambda_{i}^{+}P_{i}^{+}\textrm{ ,}\label{eq:spectrum-decomposition}
\end{equation}
we separate the positive and negative eigenvalues and order its eigenvalues
from the smallest to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}\le\ldots\le\lambda_{1}^{-}\le0\le\lambda_{1}^{+}\le\ldots\le\lambda_{N^{+}}^{+}\textrm{ .}
\end{equation}
Let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$ and $\Delta\mu^{*}\left(P_{i}^{s\prime}\right)=\mu^{*}\left(P_{i}^{s\prime}\right)-\mu^{*}\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
We can combine the classical Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with Eq.~(\ref{eq:quantum-expectation}) and define the expectation
value of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{eqnarray}
\expval{\mathbf{O}}_{\bar{\mu}} & = & \sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta\mul\left(P_{i}^{+\prime}\right),\Delta\mur\left(P_{i}^{+\prime}\right)\right]\nonumber \\
 &  & +\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta\mul\left(P_{i}^{-\prime}\right),\Delta\mur\left(P_{i}^{-\prime}\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{eqnarray}

@@

\noindent Having these two definitions, we can write down our idea
formally as follow.
